Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-04-29 _tex/paper2.tex
Built site for gh-pages

Finally, as discussed in the previous section, removing \emph{misfit}
judges based risks discarding valuable information and even introduce
bias into the trait estimates
\citep[chap.~12]{Zimmerman_1994, McElreath_2020}. The direction and
magnitude of these biases remain unpredictable because they depend on
which judges researchers exclude from the analysis. \emph{Misfit judges}
are those whose evaluations deviate substantially from the shared
consensus due to the unique characteristics of either the stimuli or the
judges themselves
\citep{Pollitt_2012a, Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018}.

@@

In line with Thurstone's theory, SCM~\ref{fig-cj03_scm} depicts the
texts' discriminal processes \((T_{ia}, T_{hb})\) and their discriminal
difference \((D_{iahbjk})\) (see Section~\ref{sec-thurstone_theory}).
Additionally, building on the arguments developed in
Section~\ref{sec-theory-issue1b}, the SCM incorporates a key design
features of CJ experiments: the judges' biases \((B_{kj})\). These
variables, along with the outcome \((O_{iahbjk})\), constitute the
preliminary set of endogenous variables,
\(V = \{ O_{iahbjk}, D_{iahbjk}, T_{ia}, T_{hb}, B_{kj} \}\). Finally,
the SCM presents the preliminary set of structural equations,
\(F = \{ f_{O}, f_{D} \}\), which define the non-parametric dependencies
among these variables.

>>>>

Finally, as discussed in the previous section, researchers risk
discarding valuable information and introducing bias into trait
estimates when they remove \emph{misfit} judges--that is, judges whose
assessments deviate markedly from the shared consensus
\citep{Pollitt_2012a, Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018},
considered outliers under the BTL model \citep{Grubbs_1969}. The
direction and magnitude of these biases remain unpredictable, as they
depend on which judges researchers exclude from the analysis
\citep{Zimmerman_1994, OHagan_2018, McElreath_2020}.

@@

In line with Thurstone's theory, SCM~\ref{fig-cj03_scm} depicts the
texts' discriminal processes \((T_{ia}, T_{hb})\) and their discriminal
difference \((D_{iahbjk})\) (see Section~\ref{sec-thurstone_theory}).
Additionally, the SCM incorporates a key design feature of CJ
experiments: the judges' biases \((B_{kj})\). This extension builds on
the arguments presented in Section~\ref{sec-theory-issue1b}, contending
that the discriminal difference becomes an observable outcome only
through judges' perceptions. Given that such perceptions may be
imperfect--and that each judge may carry some degree of bias
\citep[see][]{Pollitt_et_al_2003, vanDaal_et_al_2016}--it is reasonable
that judges' perceptions (bias) should be treated as an integral
component of the CJ system from the outset, as this leads to a more
accurate representation of the data-generating process underlying the
pairwise comparisons.

Notably, all these variables and the outcome \((O_{iahbjk})\) constitute
the preliminary set of endogenous variables,
\(V = \{ O_{iahbjk}, D_{iahbjk}, T_{ia}, T_{hb}, B_{kj} \}\). Finally,
the SCM presents the preliminary set of structural equations,
\(F = \{ f_{O}, f_{D} \}\), which define the non-parametric dependencies
among these variables.
